---
title: "Identifying relevant groups with rank functions"
author: "Carlos Morales-Torrado"
date: "2021-06-30"
tags:
- rank
- profiles
categories: R
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8,
  fig.height=8
)
```

## Introduction

This article introduces the **rank functions** that are included in the [**wpa** R package](https://microsoft.github.io/wpa). These are powerful functions that scan a full person query dataset and automatically identify groups of interest, i.e. those exhibiting the highest or lowest values for a metric in the population.

Rank functions can uncover previously unknown patterns and interesting trends, without analysts having to manually slice the data across multiple attributes. You can think of them as a tool for _automatic group exploration_ - an efficient method for identifying the top and bottom performers in a group of employees without manually reviewing all the possible organisational divisions.

Additionally, these functions can reveal what combination of attributes best describe a group that shows a specific behaviour (i.e for example, what organisational attributes represent employees suffering from excessive workload, receiving insufficient coaching, etc). As business leaders often want to develop targeted interventions to address specific issues, these functions can be a great way to inform action planning and transformation initiatives.

This article is divided in four sections. After this introduction, section two describes rank functions and show how them can be used with person query outputs. In the following section, we explore their use for profiling - i.e. to identify in which groups are specific behaviour is most common. The forth and final section, includes a list of additional use cases that can benefit from this method. 



## Rank Functions

### Baseline analysis

To get started, let's open a new session in R and load the [**wpa** R package](https://microsoft.github.io/wpa). We will also use **dplyr** to facilitate data manipulation:

```{r message=FALSE, warning=FALSE}
library(wpa)
library(dplyr)
```

Using the `sq_data` demo dataset (that contains a weekly person query output) let's use the `collaboration_rank()` function:

```{r message=FALSE, warning=FALSE}

sq_data %>% 
    collaboration_rank(return = "table")

```

The output table contains the results from the automatic group exploration. The rank function has taken all available organisational attributes in this dataset (`FunctionType`, `Organization`, `LevelDesignation`, and `IsActive`) and calculated the average collaboration hours for each group within each organisational attribute. The table shows the analysed results for all groups, ranked from highest to lowest. 

To view the entire table, we can add the function `print` to the command above. 

```{r message=FALSE, warning=FALSE}
sq_data %>% 
    collaboration_rank(return = "table") %>%  
    print(n = 30)
```

This analysis shows us that the 82 employees labeled as **"FunctionType = Sales"** have the highest collaboration hours, with a weekly average of 31 hours. In second place, there are 60 individuals in **"Organization = Inventory Management"** with 26.2 hours. On the lower end, the 75 employees labelled as **"Organization = Biz Dev"** appear to have the lowest collaboration time.

This output is useful to identify groups that are worth investigating (e.g. employees labeled as `"FunctionType = Sales"`). It also reveals that splits by `FunctionType` and `Organization` are more interesting than those by `LevelDesignation`, as they have more variation; all groups in `LevelDesignation` have similar levels of collaboration, between 18 and 22 hours a week. 

Rank functions can also display this information in visual form. We can get that by changing the `return` argument from `"table"` to `"plot"`:

```{r message=FALSE, warning=FALSE}
sq_data %>% 
    collaboration_rank(return = "plot")
```

This plot contains the same information as the table above. Each row corresponds to a different organisational attribute and the bubbles show how the population is split across different subgroups (bubble size is proportional to the number of individuals). The five groups in orange are clearly the ones spending most time in collaboration, with Sales way ahead in the front. Keep in mind that is row is a different way to slice the same population, so groups across rows are not mutually exclusive. 

### Combining two HR variables

In our baseline analysis we uncovered that the 82 employees labeled as  **"FunctionType = Sales" ** have the highest average collaboration hours. However, 
it is likely that some employees in this group are more affected than others. Furthermore, there may be pockets in other parts of this company that may
 be experiencing similar levels of collaboration ovearload, but that are hidden given the  structure of the organisational data (for example, the managers within a particular business unit).

The rank includes an additional mode called "combine", that allows us to take our analysis one step further:

```{r message=FALSE, warning=FALSE}
sq_data %>% 
    collaboration_rank(return = "table", mode = "combine")
```

The output table contains similar results from the group exploration above, but with more granularity. The function has automatically created smaller groups by combining two distinct groups from the organisational attributes in scope. In this case, it seems that the 28 managers in Sales are the worst affected group, with a weekly average of 33.7 hours in collaboration per week (that sounds like a lot of meetings!). 


### Additional functions 

So far, we have limited our analysis to `collaboration_rank()`. However, the [**wpa** R package](https://microsoft.github.io/wpa) includes many other rank functions for key workplace analytics metrics:

* `afterhours_rank()`

* `email_rank()`

* `meeting_rank()`

* `one2one_rank()`

* `workloads_rank()`


Additionaly, you can use the more general version below, if you want to apply this to any other numeric variable of your choice:

* `create_rank()`


## Using rank functions for profiling

In the previous section, we explored how rank functions can help analysts uncover interesting insights, without the need of manually slicing the data across multiple attributes. 

Rank functions have an additional use: they can help us identify what combination of attributes best describe a group that shows a given behaviour. For example, we may be concerned about individuals that are experiencing both a heavy collaboration load (more than 20 hours a week on average) and long workweeks (workweek span of over 60 hours), and would like to know where in the organisations these individuals sit. 

To answer that question, let's create a new numeric custom metric (i.e. a 1/0 flag) and then use the `create_rank` function:

```{r message=FALSE, warning=FALSE}
sq_data %>%
    mutate(At_risk = ifelse(Collaboration_hours > 20 &  Workweek_span > 60 , 1, 0)) %>%
    create_rank(metric = "At_risk", mode = "combine")
```

The results of this analysis, reveal that the combination of heavy collaboration load  plus long workweeks is most prevalent among the 13 employees in  **FunctionType = Sales** and **LevelDesignation = Junior IC** (the average employee of this groups, experiences high collaboration load and long workweeks 31% of the time). 


## Conclusion, use cases and other applications

In this articule, we have shown how **rank functions** can be a powerful tool to both uncover previously unknown patterns and to profile the groups that are experiencing a certain behaviour. In the context  of Workplace Analytics, we see a wide range of applications in the areas of wellbeing, ways of working, and sales excellence. Examples include: 

* **Workload:** Identifying groups with excessive collaboration and long workdays
* **Freedom to disconnect:**  Identifying groups with always-on behaviours
* **Coaching:**  Identifying groups with insufficient coaching and manager time
* **Exposure:**  Identifying groups with insufficient skip-level exposure
* **Isolation:**  Identifying groups with insufficient cross-team collaboration.
* **Sales performance:**  Identifying groups with insufficient external collaboration.


## Feedback

Hope you found this useful! If you have any suggestions or feedback, please visit https://github.com/microsoft/wpa/issues.
